\chapter{Nepal Attribute Distribution Justification}
\label{app:e_distributions}

\section*{Payment Channel Distribution}

The payment channel attribute reflects Nepal's digital
payment infrastructure as of 2024--2025.

\begin{table}[H]
\centering
\caption{Payment channel distribution --- Nepal context}
\begin{tabular}{clcc}
\toprule
\textbf{Value} & \textbf{Channel} &
\textbf{Probability} & \textbf{Justification}\\
\midrule
0 & Mobile wallet  & 0.60 &
eSewa and Khalti market dominance\\
1 & QR code        & 0.25 &
Rapid QR adoption at merchants\\
2 & NFC            & 0.10 &
Urban smartphone users only\\
3 & Offline token  & 0.05 &
Rural low-connectivity deployment\\
\bottomrule
\end{tabular}
\end{table}

\section*{Transaction Zone Distribution}

The transaction zone attribute reflects Nepal's geographic
population distribution across five distinct zones.

\begin{table}[H]
\centering
\caption{Transaction zone distribution --- Nepal context}
\begin{tabular}{clcc}
\toprule
\textbf{Value} & \textbf{Zone} &
\textbf{Probability} & \textbf{Justification}\\
\midrule
0 & Kathmandu valley & 0.35 &
Capital digital activity concentration\\
1 & Hill urban       & 0.20 &
Secondary cities --- Pokhara, Biratnagar\\
2 & Hill rural       & 0.25 &
Largest rural population segment\\
3 & Terai urban      & 0.12 &
Southern plains urban centres\\
4 & Terai rural      & 0.08 &
Southern plains rural population\\
\bottomrule
\end{tabular}
\end{table}

\section*{Excluded Attribute: Remittance Corridor}

Remittance corridor was considered as a third Nepal-specific
attribute but excluded on scope grounds. This thesis models
a domestic CBDC under NRB jurisdiction. Cross-border
remittances involve separate settlement rails, foreign
correspondent banking relationships, and regulatory
frameworks outside NRB's domestic CBDC mandate. Remittance
corridor is not an attribute observable within the domestic
CBDC transaction record.

Had remittance corridor been included, it would have taken
five values: domestic (55\%), India corridor (20\%), Gulf
corridor (15\%), Malaysia corridor (7\%), and other (3\%),
based on Nepal Rastra Bank remittance flow data. Its skewed
distribution would likely have produced
$s(\text{remit}) \approx 0$ at $N = 1{,}000$, decreasing
$R_{\mathrm{norm}}$ for the same reason the included
Nepal-specific attributes do.