\newpage\thispagestyle{empty}
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\vspace{0.5cm}

I would like to express my sincere gratitude to my supervisors,
\textbf{Dr. Gokul KC and Dr. Samir Shrestha}, Associate Professors,
Department of Mathematics, Kathmandu University, for the invaluable
guidance, constructive feedback, and consistent encouragement
provided throughout the course of this research.

I am grateful to the faculty members of the Department of Mathematics,
School of Science, Kathmandu University, for the academic
foundation and the learning environment that made this research
possible.

I would also like to thank my family, for their patience,
understanding, and unwavering support throughout the duration
of this programme. I would like to extend my thanks to my brother for guiding through this project.

Nepal is at a turning point in its financial history --- millions
of citizens who have never held a bank account may soon transact
digitally for the first time. That opportunity is real, but so is
the risk that the systems built to serve them could expose their
lives in ways they never consented to. I chose this topic because
privacy in a CBDC is not a technical detail --- it is a question
of whether the people who benefit least from the digital economy
will also bear the greatest privacy cost. I hope this work gives
NRB one more tool to ensure that does not happen.

\vspace{2cm}

\textbf{Prashish Dahal} \newline
\textit{April, 2026} \newline
\textit{Kathmandu, Nepal}